Dark energy is due to a math error from 1930

### 2024-12-14 paper/paper.tex
Mention the 1930 paper

\hypersetup{
pdftitle={Dark energy is due to a math error from 1935},
pdfsubject={astro-ph.CO},
pdfauthor={Logan P.~Evans},
pdfkeywords={cosmological parameters, dark energy},
}

@@

\begin{abstract}
  Observations of distant objects, such as supernovae and galaxies, is dimmed
  due to three phenomena associated with redshift: energy levels per photon,
  spectral bandwidth, and time dilation. Observed magnitudes are reported with
  k-corrections that should correct for these dimming effects, but since 1935,
  the equations used to perform these k-corrections have been missing the
  correction factor for time dilation. When observations of Type Ia supernova
  are corrected for this missing factor, the relationship between luminosity
  distance and redshift for Type Ia supernova becomes linear. This indicates
  that the expansion rate of the universe is not accelerating and there is no
  need for dark energy to explain observational data.
\end{abstract}

@@

Observations of Type Ia supernova are fundamental to the study of cosmology.
These measurements are necessary to calibrate models that measure redshift and
estimate distance. This relationship, often called the Hubble-Lema\^{i}tre law,
[TODO: find citation] describes how quickly the universe is expanding. However,
\citet{riess1998} and \citet{perlmutter1999} presented evidence that there
isn't a linear relationship between redshift and distance, but instead, distant
objects are father away than their redshift would predict (see Figure
\ref{fig:mu_distance_vs_redshift}). This phenomenon, whatever its source, is
referred to as dark energy.

@@

It turns out that measuring the apperant brightness of a Type Ia supernova is
non-trivial. The aspect of the process that we will explore here concerns how
redshift affects the light we observe. This problem is often referred to as
k-corrections, and one of the first mathematical treatments of the problem was
performed by \citet{desitter1934}.

Willem de Sitter recognized three issues that would reduce the observed
magnitude of a distant observation. First, a redshifted photon carries less
energy than when it was originally emitted. Second, a narrow spectrum of
emitted light will be spread out over a larger observed spectrum. Finally, time
dilation will reduce the rate at which an observer detects photons. The
correction for each of these issues is identical: take a measurement and
multiply it by the factor $1 + z$.

A year later, \citet{hubble1935} published a similar set of calculations for
k-corrections, but these equations used $(1 + z)^2$ instead of the $(1 + z)^3$
correction term used by de Sitter. The Hubble k-corrections have been used ever
since.

Hubble and de Sitter were clearly aware of all three dimming effects associated
with redshift, but by \citet{oke1968energy}, the two factors of $(1 + z)$ were
attributed to the change in energy and to the spectral bandwidth elongation,
which leaves time dilation as the factor that was omitted.

In 1935, the correction of $(1 + z)^2$ empirically lined up the data so that
there seemed to be a linear relationship between redshift and distance. It
wasn't until \citet{riess1998}, sixty-three years later, that the omission of a
third $(1 + z)$ correction factor resulted in faulty conclusions.

@@

The supernova data collected by the DES Collaboration used a CCD camera, a
photon counting device, as described by \citet{flaughter2015}.  \citet{kim1996}
noted that photometric measurements that depend on bolometers will need to be
corrected for the reduced energy level of redshifted light, but with a photon
counting device, this correction is not necessary.

@@

However, as shown by \citet{blonden2008} and \citet{white2024}, distant
supernova do experience time dilation. Based on this, we can reject the tired
light hypothesis and assume the big bang model.

@@

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{mu_distance_vs_redshift.png}
  \caption{The relationship between distance and redshift for two treatments of
  magnitude data. The displayed points are roughly a third of the values in the
  full DES dataset, selected evenly to aid visibility. The $k(z) = 1$ treatment
  is clearly non-linear while the $k(z) = 1 + z$ treatment appears to be
  linear.}
  \label{fig:mu_distance_vs_redshift}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{k-corrections_for_photon_counts.png}
  \caption{An example of how k-corrections take an observed spectrum and
  produce a rest-frame spectrum. [TODO]
  }
  \label{fig:mu_distance_vs_redshift}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{velocity_vs_distance.png}
  \caption{The relationship between expansion velocity and distance. The slope of this graph demonstrates the Hubble constant.
  }
  \label{fig:mu_distance_vs_redshift}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{bootstrapped_H0.png}
  \caption{A histogram of 100000 bootstrap trials measuring the Hubble constant
  $H0$. Each trial samples an absolute Type Ia magnitude $M \sim Norm(-19.2334,
  0.0404)$ based on data published by \citet{camarena2020}. It then samples,
  with replacement, a population of supernovae from the dataset published by
  \citet{abbott2024}. Finally, it uses the non-parametric linear regression
  technique described by \citet{siegel1982}. The result of the bootstrap is $H0
  \sim Norm(65.94, 1.29)$.
  }
  \label{fig:mu_distance_vs_redshift}
\end{figure}

>>>>

\hypersetup{
pdftitle={Dark energy is due to a math error from 1930},
pdfsubject={astro-ph.CO},
pdfauthor={Logan P.~Evans},
pdfkeywords={cosmological parameters, dark energy},
}

@@

\begin{abstract}
  Light from distant objects, such as supernovae and galaxies, is dimmed
  due to three phenomena associated with redshift: energy levels per photon,
  spectral bandwidth, and time dilation. Observed magnitudes are reported with
  k-corrections that should correct for these dimming effects, but since 1930,
  the equations used to perform these k-corrections have been missing the
  correction factor for time dilation. When observations of Type Ia supernova
  are corrected for this missing factor, the relationship between luminosity
  distance and redshift for Type Ia supernova becomes linear. This indicates
  that the expansion rate of the universe is not accelerating and there is no
  need for dark energy to explain observational data.
\end{abstract}

@@

Observations of Type Ia supernova are fundamental to the study of cosmology.
These measurements are necessary to calibrate models that measure redshift and
estimate distance. This relationship, often called the Hubble-Lema\^{i}tre law,
[TODO: find citation] describes how quickly the universe is expanding. However,
\citet{riess1998} and \citet{perlmutter1999} presented evidence that there
isn't a linear relationship between redshift and distance, but instead, distant
objects are farther away than their redshift would predict (see Figure
\ref{fig:mu_distance_vs_redshift}). This phenomenon is referred to as dark
energy.

@@

It turns out that measuring the apperant brightness of a Type Ia supernova is
non-trivial. The aspect of the process that we will explore here concerns how
redshift affects the light we observe. This problem is often referred to as
k-corrections, and one of the first mathematical treatments of the problem was
performed by \citet{tolman1930}. However, when Tolman made his derivation, he
did not consider the effects of a spectra that is stretched out due to
redshift.

A few years later, \citet{desitter1934}, discussed all three issues that would
reduce the observed magnitude of a distant observation. The correction for each
of these issues is identical: take a measurement for luminosity and multiply it
by the factor $1 + z$.

A year later, \citet{hubble1935} published a similar set of calculations for
k-corrections, but these equations used $(1 + z)^2$ instead of the $(1 + z)^3$
correction term used by de Sitter. They started their derivation by copying the
incorrect equation from 1930. After their derivation, they noted,

\begin{quote}
It should be specially noted that this expression differs from the correction
to m proposed by de Sitter, which contains the term $(1 + z)^3$ instead of
$(1 + z)^2$. Expression (28), however, would seem to give the proper correction
to use in connection with our equation (21), since it has been derived in such
a way as to make appropriate allowance, first, for the double effect of nebular
recession in reducing both the individual energy and the rate of arrival of
photons, and then for the further circumstance that a change in spectral
distribution of the energy that does arrive will lead to changes in its
photographic effectiveness.
\end{quote}

The Hubble k-corrections with the incorrect correction term have been used ever
since.

By \citet{oke1968}, the two factors of $(1 + z)$ were attributed to the change
in energy and to the spectral bandwidth elongation, which leaves time dilation
as the factor that was omitted.

In 1935, the correction of $(1 + z)^2$ empirically lined up the data so that
there seemed to be a linear relationship between redshift and distance. It
wasn't until \citet{riess1998}, sixty-eight years later, that the omission of a
third $(1 + z)$ correction factor resulted in faulty conclusions.

@@

The supernova data collected by the DES Collaboration used a CCD camera, a
photon counting device, as described by \citet{flaugher2015}.  \citet{kim1996}
noted that photometric measurements that depend on bolometers will need to be
corrected for the reduced energy level of redshifted light, but with a photon
counting device, this correction is not necessary.

@@

However, as shown by \citet{blondin2008} and \citet{white2024}, distant
supernova do experience time dilation. Based on this, we can reject the tired
light hypothesis and assume the big bang model.

@@

\begin{figure}[ht]
  \includegraphics[width=\linewidth]{mu_distance_vs_redshift.png}
  \caption{The relationship between distance and redshift for two treatments of
  magnitude data. The displayed points are roughly a third of the values in the
  full DES dataset, selected evenly to aid visibility. The $k(z) = 1$ treatment
  is clearly non-linear while the $k(z) = 1 + z$ treatment appears to be
  linear.}
  \label{fig:mu_distance_vs_redshift}
\end{figure}

\begin{figure}[ht]
  \includegraphics[width=\linewidth]{k-corrections_for_photon_counts.png}
  \caption{An example of how k-corrections take an observed spectrum and
  produce a rest-frame spectrum. [TODO]
  }
\end{figure}

\begin{figure}[ht]
  \includegraphics[width=\linewidth]{velocity_vs_distance.png}
  \caption{The relationship between expansion velocity and distance. The slope of this graph demonstrates the Hubble constant.
  }
\end{figure}

\begin{figure}[ht]
  \includegraphics[width=\linewidth]{bootstrapped_H0.png}
  \caption{A histogram of 100000 bootstrap trials measuring the Hubble constant
  $H0$. Each trial samples an absolute Type Ia magnitude $M \sim Norm(-19.2334,
  0.0404)$ based on data published by \citet{camarena2020}. It then samples,
  with replacement, a population of supernovae from the dataset published by
  \citet{abbott2024}. Finally, it uses the non-parametric linear regression
  technique described by \citet{siegel1982}. The result of the bootstrap is $H0
  \sim Norm(65.94, 1.29)$.
  }
\end{figure}
